---
title: "Getting started with AKME"
author: "Clint Noack"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with AKME}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

This vignette walks through the AKME workflow end-to-end on synthetic
data: an adjusted Kaplan–Meier estimator of the survival function for
right-censored concentration data, plus the weighted log-rank test of
Singh et al. (2014).

## The estimator

For an observation (reported concentration) $x_i$ the adjusted
Kaplan–Meier estimator is

$$
\widehat{S}(x_i) \;=\; \prod_{x < x_i}\!\left(1 - \frac{d^w_i}{Y^w_i}\right),
$$

where
$d^w_i = \sum_{x < x_i} w_i \delta_i$,
$Y^w_i = \sum_{x < x_i} w_i$,
$w_{ik} = 1 / n_k$ is the per-observation weight at site $k$
(with $n_k$ samples at that site), and $\delta_i$ is a binary
detection indicator (0 for below-detection-limit measurements).

## Generate synthetic data

We draw three independent log-normal datasets with different means,
randomly assign each measurement to one of ten sites with uneven
probability, and simulate a mixed-detection-limit assay that produces
right-censored values.

```{r data}
library(AKME)
library(dplyr)
library(tidyr)

means <- c(1, 1.5, 2.5)
N <- 100

set.seed(8675309)
R <- data.frame(vapply(means, function(mu) rlnorm(N, mu, 1), numeric(N)))
colnames(R) <- c("Control", "High", "Highest")

R.g <- R |>
  as_tibble() |>
  pivot_longer(everything(), names_to = "Dataset", values_to = "Concentration") |>
  mutate(
    DL = sample(c(1, 10), size = n(), replace = TRUE, prob = c(0.6, 0.4)),
    Censored = as.integer(Concentration < DL),
    Concentration = if_else(Censored == 1L, DL, Concentration),
    Site = sample(LETTERS[1:10], size = n(), replace = TRUE,
                  prob = 1:10 / sum(1:10))
  ) |>
  select(Dataset, Concentration, Censored, Site)

R.g
```

## Compare two groups with a weighted log-rank test

```{r logrank}
CvsH <- filter(R.g, Dataset != "High")

CvsH_test <- log_rank(
  CvsH,
  comp_group  = "Highest",
  rho         = 1,
  method      = "sim",
  boots       = 500,
  alternative = "greater"
)

CvsH_test$G_test
CvsH_test$p.val
```

## Plot survival curves and the null distribution

`plot_wKM()` and `G_rho_hist()` return ggplot objects; in a vignette
or REPL they auto-print.

```{r plots}
grp_KM <- group_km(CvsH)
plot_wKM(grp_KM, log_scale = TRUE)

G_rho_hist(CvsH_test)
```

## Quantile estimates per group

```{r quantiles}
grp_quantiles(R.g)
```

## References

- Xie, J., & Liu, C. (2005). Adjusted Kaplan-Meier estimator and
  log-rank test with inverse probability of treatment weighting for
  survival data. *Statistics in Medicine*, 24(20), 3089–3110.
- Singh, R., Mukhopadhyay, K., & Mantel, N. (2014). Weighted log-rank
  tests with applications to censored data.
