---
title: "AKME example with dummy data"
author: "Clint Noack"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
We will use dummy data, drawn from log-normal distributions, to demonstrate the adjusted Kaplan-Meier estimator and the weighted log-rank test. This analysis requires that all relevant functions be loaded into R. This may change if the code can be successfully bundled into a package.

```{r dummy_data}
# library(devtools)
# install_github('cwnoack/weighted_kaplan_meier/AKME')
library(AKME)

# Data are drawn from three separate log-normal distributions
# Group means
means <- c(1,1.5,2.5)

# Samples in each group
N <- 100

# Generate random data
set.seed(8675309)
R <- data.frame(sapply(means, function(mu) rlnorm(N, mu, 1)))
colnames(R) <- c('Control','Test.1','Test.2')

# `mutate` adds censoring and site ID (sampled with uneven probability)
R.g <- gather(R, Dataset,Concentration,Control:Test.2) %>%
         # Assume that 60% of samples were analyzed by method with DL = 1 ppb,
         # 40% with DL = 10 ppb
  mutate(DL = sample(c(1, 10), size = nlevels(Dataset)*N,
                     replace = T, prob = c(0.6,0.4)),
         # Flag censored values and store at detection limit
         Censored = ifelse(Concentration < DL, 1, 0),
         Concentration = ifelse(Censored == 1, DL, Concentration),
         # Randomly assign to sites, higher number sites have more samples
         Site = sample(LETTERS[1:10], nlevels(Dataset)*N, replace = T,
                       prob = 1:10/sum(1:10))) %>%
  select(Dataset, Concentration, Censored, Site)

tbl_df(R.g)
```

The remainder of this analysis makes use of the adjusted Kaplan-Meier estimator for observation (reported concentration) $x_i$:
$$
\hat{S}(x_i) = \prod_{x < x_i}\left(1 - \frac{d^w_i}{Y^w_i}\right)
$$

Where $d^w_i = \sum_{x < x_i} w_i\delta_i$, $Y^w_i = \sum_{x < x_i} w_i$, $w_{ik} = \frac{1}{n_k}$, and $\delta_i$ is a binary response for a censored value (0 for BDL measurements).

```{r summarize_dummy_dat}
plot_wKM(group_km(R.g), n_ticks = 4, units = 'mcg/m3')
grp_quantiles(R.g)
# R.g %>% ddply(.(Dataset), function(df){
#   Surv_quantile(censored_data = select(df, -Dataset), type = 'PiR')
#   }) %>% spread(Dataset, Xh) %>% arrange(desc(Percentile))
```
